### 2024-09-23 paper/paper.qmd
first commit

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Rohan Alexander
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(tinytex)
library(dplyr)
library(tibble)
library(opendatatoronto)
library(janitor)
library(ggplot2)
library(knitr)

data <- read_csv(here::here("data/analysis_data/deaths_shelter_analysis_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

\textcolor{red}{total number victims}: the total number of male and female victims